---
title: "Solo Project"
format: html
---

```{r}
library(tidyverse)
library(sf)
library(viridisLite)
burritos <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2023/2023-05-16/tornados.csv')
```


```{r}
shape <- read_sf("tl_2024_us_county\\tl_2024_us_county.shp")

ohio_shape <- shape|>
  filter(STATEFP == 39)|> #Ohio state number
  mutate(COUNTYFP = as.numeric(COUNTYFP))
```

```{r}
OHburritos <- burritos|>
  filter(st == "OH", ns == 1)|> #Filtering out Ohio tornado. Number of state affected filtered to 1.
  select(om, date, loss, f1, f2, f3, f4)|>
  pivot_longer(cols = c(f1, f2, f3, f4), names_to = "county_order", values_to = "county_code")|> #pivot long so I can count
  count(county_code, name = "n_burritos")|>
  filter(county_code != 0)|> #0 is placeholder for NA so filter
  mutate(county_code = as.numeric(county_code))
```

```{r}
map_burrito <- ohio_shape |>
  left_join(OHburritos, by = c("COUNTYFP" = "county_code"))
```


```{r}
OHpop <- read_csv("PopulationEstimates.csv")
OHpop <- OHpop |>
  filter(State == "OH", Attribute == "POP_ESTIMATE_2023", FIPStxt != 39000)|>
  mutate(FIPStxt = as.numeric(str_sub(FIPStxt, 3)))
```

```{r}
map_ohpop <- ohio_shape |>
  left_join(OHpop, by = c("COUNTYFP" = "FIPStxt"))
```

```{r}
#| fig-cap: "Franklin, Van Vert, Lorain, Huron, Medina, Clinton, Miami, and Clark counties are most heavily affected."
#| fig-alt: "Franklin, Van Vert, Lorain, Huron, Medina, Clinton, Miami, and Clark counties are most heavily affected."
ggplot() + geom_sf(data = map_burrito, aes(fill = n_burritos)) + labs(title = "Ohio Counties by Number of Tornadoes, 1950 - 2022", fill = "Tornado Count", caption = "Visualized by: Benjamin Weng\nSource: TidyTuesday/NOAA") + theme_minimal() + scale_fill_viridis_c(option = "rocket", direction = -1)
```

```{r}
#| fig-cap: "Ohio population by county. Cuyahoga, Franklin, and Hamilton counties are most populated due to cities."
#| fig-alt: "Ohio population by county. Cuyahoga, Franklin, and Hamilton counties are most populated due to cities."
ggplot() + geom_sf(data = map_ohpop, aes(fill = Value)) + labs(title = "Ohio Counties by Population, 2023", fill = "Population", caption = "Visualized by: Benjamin Weng\nSource: USDA") + theme_minimal() + scale_fill_viridis_c(option = "rocket", direction = -1)
```

It almost seems like the surrounding counties of major cities have far more tornadoes. My hypotheses are:
1. Tornado report is heavily relied on sightings. Without sightings, the rural area has more tornadoes unreported.
2. Major cities monitor their tornadoes more heavily due to safety concerns. This includes monitoring surrounding counties.
3. The urban counties themselves don't seem to have too many tornadoes except for Franklin. It might be due to the nature of urban landscape is less prone to tornadoes.
